Environmental Market Design for Large-Scale Marine Conservation

### 2019-11-13 docs/Markets_Conservation_revised_Nov13.tex
Added new version of the MS

\title{Well-Designed Environmental Markets Enable Large-Scale Marine Conservation}

@@

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
While it is commonly agreed that marine conservation should expand dramatically around the world, most countries have yet to implement large-scale Marine Protected Areas (MPAs) in their own waters. When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While this may be more-than-made-up-for through spillover, benefits from large-scale MPAs typically accrue to other countries or the high seas. We show that careful design of cross-country rights-based approaches to fisheries management can actually incentivize, rather than hinder, large scale marine conservation. By combining a model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that cross-country transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific Island Nations that collectively manages tuna purse seine fishing in its members' waters\cite{havice_2013,aqorau_2018} (Fig. \ref{fig:PNA_map}). These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP \cite{ffa_2017b}. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by December 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how a market can be designed to incentivize conservation.

Not all market-based approaches to environmental management contain the same conservation incentives. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these design features of a fisheries market affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure would surely benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could trade the rights to that spillover to adjacent countries, this could offset the foregone fisheries revenue. But if the conserving country was not allowed to trade these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior. In the following section, we will show how trading and allocation rules can shape the incentives that may facilitate or hinder large-scale conservation. We will then make use of vessel-tracking data to analyze a real-world case of large-scale marine conservation under fishing effort markets.

@@

We have shown that trading significantly reduces the costs of conservation. However, a new question arises. How should rights be re-allocated every year once a country starts closing its waters to fishing? Customarily, the allocation of fishing effort rights is a formula that combines historical fishing effort and biomass in a country's waters \cite{havice_2013}. We test a range of allocation rules that weigh effort ($\alpha$) and biomass ($1 - \alpha$) differently as the basis for ongoing rights allocation. We simulate a fishery operating with closures for 50 years and compare the resulting revenues to a fishery without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in long-term losses to the conserving country of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, a biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in revenue losses as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation. This result implies that if allocation is based purely on the biological features of a stock (\emph{e.g.} the biomass within a nation's waters), and not on fishing intensity, the incentives for conservation can be sustained through time.

@@

@@

While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model indicates that moving from the current 40\% biomass-based allocation rule to a 100\% biomass-based rule would ensure long-term financial security in the presence of large-scale MPAs, and further incentivize conservation within the PNA. This is a model prediction that cannot be empirically tested in the PNA context because the allocation rule has not been experimentally modified. Nevertheless, the PNA countries have shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may also have become pioneers in effective large-scale marine conservation in a market-based setting.

@@

We use these data to calculate the number of vessel-days that the 318 purse seiners spent in each PNA country and in PNA waters as a whole (Fig. \ref{fig:empirical}). The vessel-day equivalent of a day of fishing depends on vessel size, a measure used to control for effort creep. The Palau Arrangement for the Management of the Western Pacific Fishery Management Scheme \cite{pna_2016} states that a day of activity by vessels smaller than 50 meters long counts as half a vessel-day, a day of activity by vessels 50 - 80 meters long counts as one vessel-day, and a day of activity by vessels larger than 80 meters counts as 1.5 vessel-days. Vessel length is an observable characteristic in our dataset, and therefore vessel-days calculated in our analyses correspond to the PNA definition of vessel-days.

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell. This process results in 14 gridded datasets of fishing hours (7 years, for two groups). For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group, and normalize the value of each grid cell by diving it by the largest within-group absolute change:

@@

@@

>>>>

\title{Environmental Market Design for Large-Scale Marine Conservation}

@@

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
It is commonly agreed that marine conservation should expand dramatically around the world. However, most countries have yet to implement large-scale no-take Marine Protected Areas (MPAs). When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While biodiversity and spillover fishing benefits may far exceed these loses, benefits from large-scale MPAs typically accrue to other countries or the high seas. To overcome this dilemma, we design and test an international fisheries management scheme with transferable fishing rights that incentivizes, rather than hinders, large-scale marine conservation. By combining a model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that cross-country transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation. Our work focuses on the Vessel-Day Scheme, an environmental market employed by the Parties to the Nauru Agreement (a group of nine Pacific Island Nations) to manage their tuna fisheries, and where large-scale conservation interventions have taken place. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific Island Nations that collectively manages tuna purse seine fishing in its members' waters\cite{havice_2013,aqorau_2018} (Fig. \ref{fig:PNA_map}). These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP \cite{ffa_2017b}. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by January 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how a market can be designed to incentivize conservation.

Not all market-based approaches to environmental management contain the same conservation incentives. A pervasive finding across a range of natural resources is that features of markets, such as the allocation of rights, can have implications for the market's functioning \cite{libecap_1989}. In the context of fishing effort markets, we find that two market design features are pivotal in determining the incentives for large-scale marine conservation: trading and allocation rules. But why would these design features of a fisheries market affect the incentives for conservation? Consider the incentives for a country to close 100\% of its waters. Such a closure might benefit other countries through the spillover of fish from the protected area to the waters of neighboring countries. If the conserving country could trade the rights to that spillover to adjacent countries, this could offset the foregone fisheries revenue. But if the conserving country was not allowed to trade these rights, then it would lose all of its fishing revenue. The rules for how fishing rights are allocated across countries are equally important. Suppose rights are allocated each year based on the previous year's fishing effort: the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior. In the following section, we will show how trading and allocation rules can shape the incentives that may facilitate or hinder large-scale conservation. We will then make use of vessel-tracking data to analyze a real-world case of large-scale marine conservation under fishing effort markets.

\section{Results}

@@

We have shown that trading significantly reduces the costs of conservation. However, a new question arises. How should rights be re-allocated every year once a country starts closing its waters to fishing? Customarily, the allocation of fishing effort rights is a formula that combines historical fishing effort and biomass in a country's waters \cite{havice_2013}. We test a range of allocation rules that weigh effort ($\alpha$) and biomass ($1 - \alpha$) differently as the basis for ongoing rights allocation. We simulate a fishery operating with closures for 50 years and compare the resulting revenues to a fishery without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in long-term losses to the conserving country of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, a biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in revenue losses as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation. This result implies that if allocation is based purely on the biomass within a nation's waters, and not on fishing intensity, the incentives for conservation can be sustained through time.

@@

\section{Discussion}

@@

While trading would allow Palau, Kiribati, and other PNA members to reduce revenue losses from large-scale conservation, our model indicates that moving from the current 40\% biomass-based allocation rule to a 100\% biomass-based rule would ensure long-term financial security in the presence of large-scale MPAs, and further incentivize conservation within the PNA. A 100\% biomass-based allocation rule means that fishing rights will be distributed between nations based on the proportion of total biomass within their waters, regardless of historical fishing intensity. Furthermore, an effort-based allocation would reward undesirable behavior by granting more fishing rights to a country that fishes the most. This is a model prediction that cannot be empirically tested in the PNA context because the allocation rule has not been experimentally modified. Nevertheless, the PNA countries have already shown that rights-based management of transboundary resources can result in large management and economic benefits \cite{havice_2013,aqorau_2018}. By facilitating trade and allocating rights based on biomass, they may also have become pioneers in effective large-scale marine conservation in a market-based setting.

@@

We use these data to calculate the number of vessel-days that the 318 purse seiners spent fishing in each PNA country and in PNA waters as a whole (Fig. \ref{fig:empirical}). The vessel-day equivalent of a day of fishing depends on vessel size, a measure used to control for effort creep. The Palau Arrangement for the Management of the Western Pacific Fishery Management Scheme \cite{pna_2016} states that a day of activity by vessels smaller than 50 meters long counts as half a vessel-day, a day of activity by vessels 50 - 80 meters long counts as one vessel-day, and a day of activity by vessels larger than 80 meters counts as 1.5 vessel-days. Vessel length is an observable characteristic in our dataset, and therefore vessel-days calculated in our analyses correspond to the PNA definition of vessel-days.

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell. We use a 1-degree grid as a reasonable compromise between higher resolutions that would result in a more granular but noisy footprint, and the simple estimation of vessel-days at the EEZ-level (as in Fig S6, but represented spatially). This process results in 14 gridded datasets of fishing hours (7 years, for two groups). For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group, and normalize the value of each grid cell by diving it by the largest within-group absolute change:

@@

\subsection{Code availability}

@@

\subsection{Competing interests}

The authors declare no competing interests.
